Progress Report for CSS 586 Project: \\ Music Generation Using Deep Learning

### 2021-05-04 papers/progress/jack/progress-jack.tex
Update to Jack's report

The original goal of the project is music transcription, which is the translation of audio data
to symbolic representation of musical notations. With further consideration, we decided to focus on 
a different problem to better align our work. The new objective is music generation using deep learning.
There are several methods that could be used to generate new music from a corpus of training data: MIDI, 
time-domain sound pressure data, frequency-domain representation of the sound pressure. Regardless of the 
data type, music generation involves the application of a sequence model such as recurrent neural network 
(RNN). The focus of my contribution is using a combination of time-domain and frequency-domain information. 
The current proges includes literature review, data exploration, learning about different representations 
of an audio signal. I am working on building a data pipeline and experiencing with RNN models.
\end{abstract}

@@

Historically, the modelling of raw audio data is extremely challenging because a single second of audio recording
could contains up to 44,100 samples in the case of music. Google DeepMind's WaveNet is one of the first models
that successfully learn from raw audio data \cite{oord_wavenet_2016}. It is designed using a technique called 
1D dialated causal convolution that allows an output to capture information from many inputs with a minimal
computational cost. WaveNet was designed for the primary purpose of generating speech which mimics any human voice.
However, since the architecture can be used to model any raw audio signal, it was also used to generate music.
Although the outputs were far from being any masterpieces, the results show the possibility of generating musical
pieces from the raw audio data.

\section{Planned Methods}

\subsection{Datasets}

>>>>

The original goal of the project is music transcription, which is the translation of audio data
to symbolic representation of musical notations. With further consideration, we decided to focus on 
a different problem to better align our work. The new objective is music generation using deep learning.
There are several methods that could be used to generate new music from a corpus of training data: MIDI, 
time-domain sound pressure data, frequency-domain representation of the sound pressure. Regardless of the 
data type, music generation involves the application of a sequence model such as recurrent neural network 
(RNN). The focus of my contribution is using a combination of time-domain and frequency-domain information. 
The current proges includes literature review, data exploration, learning about different representations 
of an audio signal. We are working on building a data pipeline and experimenting with various RNN models.
\end{abstract}

@@

Historically, the modeling of raw audio data is extremely challenging because a single second of audio recording
could contains up to 44,100 samples. Google DeepMind's WaveNet is one of the first models
that successfully learn from raw audio data \cite{oord_wavenet_2016}. It is designed using a technique called 
1D dialated causal convolution that allows an output to capture information from many inputs with a minimal
computational cost. WaveNet was designed for the primary purpose of generating speech which mimics any human voice.
Since the architecture can be used to model any raw audio signal, it was also used to generate music.
Although the outputs were far from being any masterpieces, the results show the possibility of generating musical
pieces from the raw audio data.

An alternatve representation of the time-series signal is the frequency-domain representation. Existing generative
models for audio have largely focused on the time-domain waveforms. In 2019, a generative model for audio in the
frequency domain called MelNet was introduced \cite{vasquez2019melnet}. MelNet models the spectrograms, which are
time-frequency representations of audio. The advantage of spectrograms is that they are compressed expressions of 
the time-domain signal. MelNet leverages this advantage to generate high-fidelity audio samples, which capture 
structures that are still challenging for time-domain models.

\section{Planned Work}
The focus of this project is generating short musical pieces by modeling the spectrograms of input audio. The
following subsections describe our plan toward accomplishing this objective.

\subsection{Datasets}
We use a dataset called MusicNet published by University of Washington, the department of Com[uter Science 
\cite{thickstun2017learning}. This dataset is a collection of 330 freely-licensed classical music recordings 
captured under various studio and microphone conditions. The recordings are in .wav raw audio format and are
polyphonic, meaning that multiple musical notes could present at a single point in time.
